---
title: "How long did each Australia Prime Minister lived, by year of birth?"
author: "Joseph Chung"
date: today
date-format: "MMM D, YYYY"
format:
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: Tutorial 7 topic Web Scraping
    number-sections: true
    colorlinks: true
    lof : true
    lot : true
thanks: "Code and data are available at: https://github.com/UtopianYoungChung/Tutorial-7.git"
bibliography: references.bib
---

```{r setup, include=FALSE}
library(rvest)
library(xml2)
library(tidyverse)
library(palmerpenguins)
library(knitr)
library(ggplot2)
```

# Findings

I tried to copy one of the examples from the textbook after reading about what was expected for this week's tutorial [@citeinf312]. I first attempted to scrape data from the provided Wikipedia websites using the "SelectorGadget," but I couldn't seem to get it to work. When I couldn't get it to work, I manually scraped the data from the websites using the html element (".wikitable") code. For each site, on three out of the four sites, the code extracted the data from a small table, which is unimportant for the tutorial. But luckily, one site (Australia) came back with information from a large table with the dataset we need (although we found some missing values that will be discussed later in the paper).

# Data

The data was scraped from "List of prime ministers of Australia -- Wikipedia" [@citewebsite] The scraped data contained 10 variables and 76 observations named 'parse_data_selector_gadget." From this set, we cleaned the data to our needs, and it was reduced to 4 variables and 31 observations: name, born, died, and age_at_death, as shown in @tbl-lifespan. @fig-lifespan illustrates each Prime Minister's lifespan on a graph.
\newpage
```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: tbl-lifespan
#| tbl-cap: Australian Prime Minister and their life span
pm_lifespan <- read_csv(here::here("inputs/data/cleaned_data.csv"))
pm_lifespan <- pm_lifespan %>% select(-...1)

pm_lifespan %>%
  head() %>%
  knitr::kable(
    col.names = c("Prime Minister", 
                  "Birth year", 
                  "Death year", 
                  "Age at death")
  )
```

```{r}
#| messages: false
#| echo: false
#| warning: false
#| label: fig-lifespan
#| fig-width: 8
#| fig-height: 14
#| fig-cap: Australian Prime Minister and their life span
pm_lifespan <- read_csv(here::here("inputs/data/cleaned_data.csv"))
pm_lifespan <- pm_lifespan %>% select(-...1)

pm_lifespan |>
  mutate(
    still_alive = if_else(is.na(died), "Yes", "No"),
    died = if_else(is.na(died), as.integer(2023), as.integer(as.numeric(died)))
  ) |>
  mutate(name = as_factor(name)) |>
  ggplot(aes(
    x = born,
    xend = died,
    y = name,
    yend = name,
    color = still_alive
  )) +
  geom_segment() +
  labs(
    x = "Year of birth",
    y = "Prime minister",
    color = "PM is currently alive",
    title = "How long each UK Prime Minister lived, by year of birth"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

```
\newpage

# Reflection

However, there was missing data on the Prime Ministers who were still alive; 8 observations were removed from the plot. Therefore, we are missing the information of the PMs who have yet to die; all 4 variables on the PM are missing.

With more resources and time, I propose doing some more scraping and developing codes to scrape all the information.

\newpage
 
# References
